---
title: "Data 607 - Assignment 1 - Loading Data into a Data Frame"
author: "Anna Moy"
date: "2024-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The article that I choose from fivethirtyeight.com was Where People Go To Check The Weather. The article discuss the different methods in which people would check the weather. The highest percentage of people would check the weather through the phone's default weather app. Depending on the age and location of the respondents the percentage of people checking weather would vary.

Link to the article:[Where People Go To Check The Weather](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/)

**Load Library**

```{r, load-library, message=FALSE}
library(tidyverse)
library(ggplot2)
```

**Load the .csv into a dataframe**

```{r, load-csv, message=FALSE}
weather <-  read_csv("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/weather-check.csv")
```

**Rename the columns**

```{r, rename-columns}
colnames(weather) <- c("id", "check_weather","source","web_or_app","smartwatch","age","gender","income","us_region")
```

### Bar Plot

What is the number of respondents that check the weather?

```{r plot-check-weather, echo=FALSE}
ggplot(weather, aes(check_weather, color = check_weather, fill = check_weather)) +
  geom_bar() +
  xlab("Respondents that check the weather") +
  ylab("# of respondents")
```

# Conclusion

One of my recommendation is to add additional survey questions such as -What time of day the respondents would check the weather? -What is the time of day the respondent goes outside? -The number of times a respondent checks the weather It will bring a helpful color to see if people are checking the weather before they go out.
