---
title: "Assignment5"
author: "Anna Moy"
date: "2024-02-25"
output: html_document


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-library, message = FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

### Read the csv file for airlines
```{r read-csv, message = FALSE}
airline <- read_csv("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/airline.csv")

airline
```

### Remove the blank line on data
```{r read-blank-line}
airline <- airline[-3,]
```
### Fill in blank data for airline
```{r fill-in-blank-data}
airline %>%
  fill('...1', .direction = "down")
```

### Transpose the data and rename column title
```{r transpose}
airline <- airline %>%
  fill('...1', .direction = "down") %>%
  pivot_longer(-c('...1', '...2'), names_to = "location") %>%
  pivot_wider(names_from ='...2', values_from = value) %>%
  rename("airline"= '...1', "on_time" = 'on time')
airline
```

### Find the average and median for each airline delays and on time
```{r airline-delays}
airline %>%
  group_by(airline) %>%
  select (airline, on_time, delayed) %>%
  summarize(meandelay = mean(delayed), mediandelay = median(delayed), meanontime = mean(on_time), medianontime = median(on_time))
```

### Find the percent delay for each airline by location in descending order
```{r delay-by_location}
flight <- airline %>%
  group_by(airline) %>%
  mutate(total_delay = sum(delayed), total_ontime = sum(on_time),all_flight = total_delay + total_ontime, percent_rate = delayed/all_flight, percent_delay = (delayed/all_flight)) %>%
  arrange(airline, desc(percent_delay))
flight
```

### Total flights for Alaska and AM West
```{r total-delays}
total <- airline %>%
  group_by(airline) %>%
  summarize(sum = sum(delayed)) 

total
```

### flights by location


### Summary of each airline on time and delay data
```{r summary-each-airline}
airline %>%
  filter(airline == "ALASKA") %>%
  summary(airline)

airline %>%
  filter(airline == "AM WEST") %>%
  summary(airline)
```

### Plot delay by location for each airline
```{r plot-delay}
ggplot(flight, aes(x = location, y = percent_delay, fill = airline)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r plot-total}
ggplot(total, aes(x = airline, sum, fill = airline)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r plot-delayed_percent}
ggplot(flight, aes(x = airline, percent_rate, fill = airline)) +
  geom_bar(stat = "identity", position = "dodge")
```

## Conclusion
AM West has more delays than Alaska as the average and median are higher than Alaska. But if we look at the overall flights compared to the percentage of delay we will notice that Alaska has more delays then AM West. There was a bias if we were jsut looking at over all delayed in total without looking overall picture. 