---
title: "Assignment3"
author: "Anna Moy"
date: "2024-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-tidyverse, message=FALSE}
library(tidyverse)
library(readr)
```

1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

Loading the .csv file for College Majors data
```{r select-certain-word, message = FALSE}
degree <-  read_csv("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/majors-list.csv")

degree <- degree %>%
  filter(grepl("DATA", Major) | grepl("STATISTICS", Major))

degree
```

2. Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
berry <- c('[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"')

a <-str_sub(berry, 5, 17)
b <-str_sub(berry, 20, 29)
c <-str_sub(berry, 35, 46)
d <-str_sub(berry, 50, 63)
e <-str_sub(berry, 70, 80)
f <-str_sub(berry, 85, 96)
g <-str_sub(berry, 101, 113)
h <-str_sub(berry, 115, 126)
i <-str_sub(berry, 135, 146)
j <-str_sub(berry, 150, 155)
k <-str_sub(berry, 165, 172)
l <-str_sub(berry, 180, 189)
m <-str_sub(berry, 201, 207)
n <-str_sub(berry, 216, 228)

berry2 <- str_c("c(",a, ", ",b,", ",c,", ",d,", ",e,", ",f,", ",g,", ",h,", ",i,", ",j,", ",k,", ",l,", ",m,", ",n,")")

writeLines(berry2)
```

3. Describe, in words, what the expressions will match: 
- (.)\1\1
This is three letters of the same character.
```{r}
data <- c("aaa")
str_view(data,"(.)\\1\\1")
```
- "(.)(.)\\2\\1"
It represents 2 letters(2 same letters or different) and then the letters are switched in order. 

```{r}
data <- c("abba")
str_view(data,"(.)(.)\\2\\1")
```
- (..)\1

It represents 2 letters (2 same letters or different) that are repeating.
```{r}
data <- c("baba")
str_view(data, "(..)\\1")
```

- "(.).\\1.\\1"
This represents 1 letter repeated 3 times with a letter in between the repeating letters.
```{r}
data <- c("abaca")
str_view(data, "(.).\\1.\\1")
```

- "(.)(.)(.).*\\3\\2\\1"
This represents 3 letter(same or different letters) with no letter or another random letter that can repeat and the original 3 letter switched around.
```{r}
data <- c("abceecba")
str_view(data,"(.)(.)(.).*\\3\\2\\1")
```

4. Construct regular expressions to match words that:
- Start and end with the same character.
  expression: (.).\\1
``` {r}
data <- c("TOT")
str_view(data, "(.).\\1")
```
- Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
  expression: (..)\\1
```{r}
data <- c("chch")
str_view(data, "(..)\\1")
```

- Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
  expression: (.).\\1.\\1
```{r}
data <- c("pupap")
str_view(data, "(.).\\1.\\1")
```