---
title: "Assignment 7"
author: "Anna Moy"
date: "2024-03-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-library, message = FALSE}
library(rvest)
library(xml2)
library(jsonlite)
library(tibble)
```

# Read the HTML and make it into a dataframe

```{r read-html}
books<- read_html("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/Books.html", trim = T, as.data.frame = T)

books<- books %>%
  html_node("table")%>% 
  html_table(header =TRUE, fill = TRUE) 
books

#Another way to run it for myself knowledge
# df_table <- books %>% 
#  html_element(xpath = "//table") %>% 
#  html_table()
```


# Read the XML and make into a dataframe
```{r read-xml}
booksxml <- read_xml("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/Books.xml")

title <- xml_text(xml_find_all(booksxml, xpath ="//title"))
author <- xml_text(xml_find_all(booksxml, xpath ="//author"))
genre <- xml_text(xml_find_all(booksxml, xpath ="//genre"))
published <- xml_text(xml_find_all(booksxml, xpath ="//published"))
ISBN <- xml_text(xml_find_all(booksxml, xpath ="//ISBN"))

df2<-data_frame(title,author,genre,published,ISBN)
df2
```

# Read the JSON and make into a dataframe

```{r readjson}
bookjson <- fromJSON("https://raw.githubusercontent.com/AnnaMoy/Data-607/main/Books.JSON")

json_data_frame <- as.data.frame(bookjson)
json_data_frame
```

# Conclusion

The HTML, XML and JSON data output all look fairly the same except the JSON did not print it out in the same format. They are not 100% idenitical but very similar. It was the hardest to load the XML as I had to covert the data into columns then print it out as a dataframe. 